---
title: "AA_nonGAIN_SCZ_preQC_Steps"
author: "Peter Fiorica"
date: "May 22, 2018"
output: html_document
---

1. Copy files from dbGaP directory in Wheelerlab3 to my directory.
```{bash}
#Genotypes
$cp phg000037.v1.p1.nonGAIN_Schizophrenia.genotype-calls-matrixfmt.c2.SARC.tar /home/peter/AA_nonGAIN_SCZ/
$cp phg000037.v1.p1.nonGAIN_Schizophrenia.genotype-calls-matrixfmt.c1.GRU.tar /home/peter/AA_nonGAIN_SCZ/
#Phenotypes
$cp phs000167.v1.pht000714.v1.p1.c2.nonGAIN_Schizophrenia_cases.SARC.txt.gz /home/peter/AA_nonGAIN_SCZ/
$cp phs000167.v1.pht000714.v1.p1.c1.nonGAIN_Schizophrenia_cases.GRU.txt.gz /home/peter/AA_nonGAIN_SCZ/
  #repeat this process for the pht000715 and pht000716 for both the GRU and SARC groups
```

2. Extract the contents of the .tar genotypes
```{bash}
#Be sure to view the contents of the tar folder before extracting.  You can do this using `tar -tf filename.tar.gz`
$tar -xvf phg000037.v1.p1.nonGAIN_Schizophrenia.genotype-calls-matrixfmt.c1.GRU.tar
$tar -xvf phg000037.v1.p1.nonGAIN_Schizophrenia.genotype-calls-matrixfmt.c2.SARC.tar
#Now, you have access to the bfiles for each consent group.
```

3.Confirming that their is a phenotype for every genotype:
```{bash}
zcat /home/peter/AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c1.GRU/nonGAIN_Schizophrenia_consent_GRU.fam.gz | wc -l
#2651;This is the number of people that are included in the initial genotype data for the GRU group
zcat /home/peter/AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c2.SARCnonGAIN_Schizophrenia_consent_SARC.fam.gz | wc -l
#350

##Cases for each consent group:
GRU:1274
SARC:322

#Controls_main for each consent group
GRU:1395
SARC:The table is intentionally blank because it does not include subjects for the SARC consent group 

#Controls_questionnaire for each consent group
GRU:1395
SARC:The table is intentionally blank because it does not include subjects for the SARC consent group
```

4. Isolating the AA genotype and phenotype data from the EA genotype and phenotype data:

```{r}
library(dplyr)
library(data.table)
```

#Read in data containing genotype and phenotype information for all individuals
```{r}
SARCfam<-fread("z://AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c2.SARC/nonGAIN_Schizophrenia_consent_SARC.fam",header=F)  #The .fam (genotype) file for individuals in the "Scizophrenria and Related Conditions (SARC)" consent group.
GRUfam<-fread("z://AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c1.GRU/nonGAIN_Schizophrenia_consent_GRU.fam",header =F) #The .fam (genotype) file for individuals in the "General Research Use" consent group.
SARCphenocases<-fread("Z://AA_nonGAIN_SCZ/phs000167.v1.pht000714.v1.p1.c2.nonGAIN_Schizophrenia_cases.SARC.txt",header=T) #The phenotype file confirming case for individuals in the "Scizophrenria and Related Conditions (SARC)" consent group. All individuas in the SARC group are cases.
GRUphenocases<-fread("Z://AA_nonGAIN_SCZ/phs000167.v1.pht000714.v1.p1.c1.nonGAIN_Schizophrenia_cases.GRU.txt",header=T) #The case phenotype file for individuals in the "General Research Use" consent group.
GRUphenocontrols<-fread("Z://AA_nonGAIN_SCZ/phs000167.v1.pht000715.v1.p1.c1.nonGAIN_Schizophrenia_controls_main.GRU.txt",header =T) #The control phenotype file for individuals in the "General Research Use" consent group.
```

The .fam files do not include the phenotype data (a 1 or 2 in the last column) for any of the files.  Most of the time they should.
This means that the files must be merged together by individual to ensure that each person has a phenotype; however, from looking at the information that has been read into R, it is clear that the phenotype files and the genotype files use different ID formats.  
There is a file that contains the IDs for both the genotype files and the phenotype files.  It will be used to merge the two.

```{r}
#The file that contains the IDs in NG (genotype) format and phenotype format.
AllID4Geno<-fread("z://AA_nonGAIN_SCZ/phs000167.v1.pht000719.v1.p1.nonGAIN_Schizophrenia_Sample.MULTI.txt",header =T)
naming2<-as.character(AllID4Geno[1,])
colnames(AllID4Geno)<-naming2
AllID4Geno<-AllID4Geno[-1,]
```

```{r}
GRUCaseAncestry<-GRUphenocases$phv00071341.v1.p1.c1
table(GRUCaseAncestry)
naming3<-as.character(GRUphenocases[1,])
colnames(GRUphenocases)<-naming3
GRUphenocases<-GRUphenocases[-1,]

table(SARCphenocases$phv00071341.v1.p1.c2)
naming4<-as.character(SARCphenocases[1,])
colnames(SARCphenocases)<-naming4
SARCphenocases<-SARCphenocases[-1,]

naming5<-as.character(GRUphenocontrols[1,])
colnames(GRUphenocontrols)<-naming5
GRUphenocontrols<-GRUphenocontrols[-1,]
```
#There are only 105 total self-identified African American cases in this cohort.
#21 AA individuals are controls

```{r}
#The first 20 individuals in SARCphenocase file are AA
#The first x amount of rows in each phenotype file is 
AA_SARCphenocases<-SARCphenocases[1:20,]
AA_GRUphenocases<-GRUphenocases[1:85,]
AA_GRUphenocontrols<-GRUphenocontrols[1:21,]
```

5. Identify AA cases and controls by their genotype ID. consent group.
```{r}
#Since the plink files become desynchronized when I create a .fam file from R of different individuals.  I am making a list of AA individuals to "keep" in plink.
GRUphenocase<-left_join(AA_GRUphenocases,AllID4Geno,by="dbGaP SubjID")
IDAA_GRUcases<-dplyr::select(GRUphenocase,"SAMPID_NG-format")

GRUphenocontrol<-left_join(AA_GRUphenocontrols,AllID4Geno,by="dbGaP SubjID")
IDAA_GRUcontrols<-dplyr::select(GRUphenocontrol,"SAMPID_NG-format")

SARCcases<-left_join(AA_SARCphenocases,AllID4Geno,by="dbGaP SubjID")
SARCfamIDs<-left_join(SARCcases,SARCfam,by=c("SAMPID_NG-format"="V1"))
IDAA_SARCcases<-select(SARCfamIDs, "SAMPID_NG-format", "V2")

GRUIDsAll<-rbind(IDAA_GRUcases,IDAA_GRUcontrols)
GRUfamIDs<-left_join(GRUIDsAll,GRUfam,by=c("SAMPID_NG-format"="V1"))
row.has.na <- apply(GRUfamIDs, 1, function(x){any(is.na(x))})
sum(row.has.na)
GRUfamIDS1<-GRUfamIDs[!row.has.na,]
GRUIDs2<-dplyr::select(GRUfamIDS1, "SAMPID_NG-format", "V2")
                       
fwrite(GRUIDs2,"Z://AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c1.GRU/AA_GRUlist.txt",col.names = F,row.names = F, sep = " ",quote = F)

fwrite(IDAA_SARCcases,"Z://AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c2.SARC/AA_SARClist.txt",col.names = F,row.names = F, sep=" ", quote= F)

```

```{bash}
plink --bfile nonGAIN_Schizophrenia_consent_GRU --keep AA_GRUlist.txt  --make-bed --out AAbfilesnopheno

plink --bfile nonGAIN_Schizophrenia_consent_SARC --keep AA_SARClist.txt  --make-bed --out AAbfilesnophenoSARC
```

```{bash}
#for some reason, all of the phenotypes are now identified as -9... This does not make too much sense so I am going to try to change this
awk -v s=1 '{print $1, $2 , $3, $4, $5, $6+11 }' AAbfilesnophenoSARC.fam > AAbfilesWphenoSARC.fam
#Since all of these individuals are in the SARC consent group, they are all cases.  I can just make their phenotype 2

plink --bed AAbfilesnophenoSARC.bed --bim AAbfilesnophenoSARC.bim --fam AAbfilesWphenoSARC.fam --make-bed --out AAbfilesWphenoSARC1
```

6. Generate new genotype files for 
```{r}

GRUfamcases<-left_join(IDAA_GRUcases,GRUfamIDS1, by="SAMPID_NG-format")
row.has.na <- apply(GRUfamcases, 1, function(x){any(is.na(x))})
sum(row.has.na)
GRUcasesfam<-GRUfamcases[!row.has.na,]
GRUcasesfam$V6=2 ##This completely identifies which values are cases

GRUfamcontrols<-left_join(IDAA_GRUcontrols,GRUfamIDS1, by="SAMPID_NG-format")
row.has.na <- apply(GRUfamcontrols, 1, function(x){any(is.na(x))})
sum(row.has.na)
GRUcontrolsfam<-GRUfamcontrols[!row.has.na,]
GRUcontrolsfam$V6=1

GRUfamfinal<-rbind(GRUcontrolsfam,GRUcasesfam)
fwrite(GRUfamfinal,"Z://AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c1.GRU/AAGRUbfilesWpheno.fam",row.names = F, col.names = F, sep = " ",quote = F)
```


```{bash}
plink --bed AAbfilesnopheno.bed --bim AAbfilesnopheno.bim --fam AAGRUbfilesWpheno.fam --make-bed --out AAbfilesWphenoGRU

plink --bfile /home/peter/AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c2.SARC/AAbfilesWphenoSARC1 --bmerge /home/peter/AA_nonGAIN_SCZ/phg000037.v1.p1.nonGAIN_Schizophrenia.matrixfmt.c1.GRU/AAbfilesWphenoGRU  --make-bed --out/home/peter/AA_nonGAIN_SCZ/QCSteps/AAnonGAIN_SCZ_start
```
