---
title: "PreQCForBipolarData"
author: "Peter Fiorica"
date: "10/11/2018"
output: html_document
---
These steps are broken down by consent group: General Research Use (GRU), Bipolar and Related Disorders (BARD), and Bipolar Disorder Only (BDO)

1A). Copy Appropriate files from wheelerlab3 directory to my directory.
62470_GAIN_DS-BD_n653/
57575_GAIN_GRU-BD_n1767/ 
62469_GAIN_REL-BD_n841/
```{bash}
#In 62470_GAIN_DS-BD_n653/
cp phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype.tar /home/peter/
cp phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype.tar /home/peter/
```

2A) Extract contents using tar:
```{bash}
#Be sure to view the contents of the tar folder before extracting.  You can do this using `tar -tf filename.tar.gz`
tar -xvf phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype.tar
#Before extracting the contents of the next file, I am going ot take a quick look these.
less -S phg000015.filtered.BDO.plink.fam.gz
zcat phg000015.filtered.BDO.plink.fam.gz | wc -l
#There are 378 individuals here, and they have 2 or 0 in their phenotype column, so they appear to all be cases.
zcat phg000015.filtered.BDO.plink.bim.gz | wc -l
#We are working with 729454 SNPs
# I also checked the README, and there are a little less than 40 individuals that need to have there phenotypes changes.  See /home/peter/bipolar_disorder/README.IMPORTANT
tar -xvf phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype.tar
#This has 264 individuals here
#All 2s and 0s in phenotype as well.
#We are working with 845814 data.
```

1B). Copy Appropriate files from wheelerlab3 directory to my directory.
```{bash}
#In 57575_GAIN_GRU-BD_n1767/
#These appear to be the General Research Use Data (GRU)
cp phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype.tar /home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/
  
cp phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype.tar /home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/
```
2B). Extract the contents of the file
```{bash}
tar -xvf phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype.tar
#1046 individuals here
#Phenotype is only 1s
#729454 SNPs
tar -xvf phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype.tar
#Phenotypes = 1s
#679 individuals
#845814 SNPs
```
1C). Copy the files to my directory
```{bash}
#62469_GAIN_REL-BD_n841/
cp phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype.tar /ho
me/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/

cp phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype.tar /ho
me/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/
```
2C). Extract the contents of the files.
```{bash}
tar -xvf phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype.tar
#675 Individuals
#729454 SNPs
#phenotype=2 and 0s
tar -xvf phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype.tar 
#153 individuals
#845814 SNPs
#Phenotype=2s and 0s
```

#Once I took a look at the Release Notes for the total Datasets, I realized that there are only 1101 total African American Individuals in the study.
#686 Controls (8 duplicates), 415 cases. Not a ton of data.

3) Reading in Information to exclude any non AA individuals
```{r}
AAinfoGRUcontrols<-fread("/home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/phs000017.v3.pht000383.v1.p1.c1.AA_GAIN_phenotype_controls_main_subset.GRU.txt", header =T, fill=T)
colnames(AAinfoGRUcontrols)<-as.character(c(AAinfoGRUcontrols[10,]))
AAinfoGRUcontrols <- AAinfoGRUcontrols[-c(1:10), ]#Removes rows 1 through 10 to clean up the data table.
table(AAinfoGRUcontrols$self_id_ancestry)
#This file contains only AA individuals.  These are all of the AA controls.
```
```{bash}
zcat phs000017.v3.pht000696.v1.p1.c2.DIGS2and3_PedData_Diagnosis.BARD.txt.gz | wc -l 
#381
zcat phs000017.v3.pht000695.v1.p1.c2.AA_GAIN4_Diagnosis.BARD.txt.gz | wc -l 
```

4) Generating a .fam file for AA controls
```{r}
GRU15fam<-fread(sprintf("zcat %s", "/home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype/phg000015.filtered.GRU.plink.fam.gz", header =F))
GRU16fam<-fread(sprintf("zcat %s", "/home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype/phg000016.filtered.GRU.plink.fam.gz", header =F))

GRUtotalfam<-bind_rows(GRU15fam,GRU16fam)
GRUtotalfam[duplicated(GRUtotalfam),] #no duplicates
GRUcontrolsfull<-left_join(AAinfoGRUcontrols,GRUtotalfam, by=c("individual_id"="V2"))
GRUcontrolsfam<-dplyr::select(GRUcontrolsfull,"V1","individual_id","V3","V4","V5","V6")
row.has.na<-apply(GRUcontrolsfam,1,function(x){any(is.na(x))})
sum(row.has.na)
GRUfamcontrolfinal<- GRUcontrolsfam[!row.has.na,]#This file contains the phenotype individuals for AA controls with V6(phenotype) being 1 #We now only have 671 AA controls.

aaGRUcontrollist<-dplyr::select(GRUfamcontrolfinal, V1, individual_id)

fwrite(aaGRUcontrollist, "/home/peter/bipolar_disorder/preQCSteps/AAcontrolslist.txt", col.names = F, row.names = F, quote = F, sep = "\t")
fwrite(GRUfamcontrolfinal, "/home/peter/bipolar_disorder/preQCSteps/AAcontrols.fam", col.names = F, row.names = F, quote = F, sep = "\t")
```

5) Generating a .fam file for the REL(BARD) and DS(BDO) concent groups:
```{r}
REL15<-fread(sprintf("zcat %s", "/home/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype/phg000015.filtered.BARD.plink.fam.gz", header =F))
REL16<-fread(sprintf("zcat %s", "/home/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype/phg000016.filtered.BARD.plink.fam.gz", header =F))
BARDtotal<-bind_rows(REL15,REL16)
BARDtotal[duplicated(BARDtotal),]#no duplicates

BDO15<-fread(sprintf("zcat %s", "/home/peter/bipolar_disorder/62470_GAIN_DS-BD_n653/phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype/phg000015.filtered.BDO.plink.fam.gz", header =F))
BDO16<-fread(sprintf("zcat %s", "/home/peter/bipolar_disorder/62470_GAIN_DS-BD_n653/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype/phg000016.filtered.BDO.plink.fam.gz", header =F))
BDOtotal<-bind_rows(BDO15,BDO16)
BDOtotal[duplicated(BDOtotal,)]#no duplicates


#Getting the phenotype files ready.
BDOdiagnosis<-fread(sprintf("zcat %s","/home/wheelerlab3/Data/dbGaP_Proj_15297_SCZ_BD/62470_GAIN_DS-BD_n653/GAIN/Bipolar/phs000017v3/p1/phs000017.v3.pht000695.v1.p1.c3.AA_GAIN4_Diagnosis.BDO.txt.gz", header =T))
a<-BDOdiagnosis[1,]
colnames(BDOdiagnosis)<-as.character(a)
BDOdiagnosis<-BDOdiagnosis[-1,-6]
table(BDOdiagnosis$Bipolar_Subject)
#There are 224 cases here
BARDdiagnosis<-fread(sprintf("zcat %s","/home/wheelerlab3/Data/dbGaP_Proj_15297_SCZ_BD/62469_GAIN_REL-BD_n841/GAIN/Bipolar/phs000017v3/p1/phs000017.v3.pht000695.v1.p1.c2.AA_GAIN4_Diagnosis.BARD.txt.gz",header = T))
b<-BARDdiagnosis[1,]
colnames(BARDdiagnosis)<-as.character(b)
BARDdiagnosis<-BARDdiagnosis[-1,-6]
table(BARDdiagnosis$Bipolar_Subject)
#There are 135 cases in here

#So 135 + 224 =359, which does not =415 number from the manifest.

#But the 150 + 265 =415 when you include the "other" individuals as cases.
case=c("Case")
BDOcases<-filter(BDOdiagnosis, BDOdiagnosis$Bipolar_Subject%in%case)
BDOCasePrefam<-left_join(BDOcases,BDOtotal,by=c("GAIN_ID"="V2"))
BDocasefam<-dplyr::select(BDOCasePrefam,"V1","GAIN_ID","V3","V4","V5","V6")
row.has.na<-apply(BDocasefam,1,function(x){any(is.na(x))})
sum(row.has.na)#Removed 10 individuals here
BDOcasefam<-BDocasefam[!row.has.na,]#This Fam file is good to go for consent group BARD
BDOcasefam<-mutate(BDOcasefam,V1,GAIN_ID,V3,V4,V5,V6=2)

BARDcases<-filter(BARDdiagnosis,BARDdiagnosis$Bipolar_Subject%in%case)
BARDCasePrefam<-left_join(BARDcases,BARDtotal,by=c("GAIN_ID"="V2"))
BARDcasefam<-dplyr::select(BARDCasePrefam,"V1","GAIN_ID","V3","V4","V5","V6")
row.has.na<-apply(BARDcasefam,1,function(x){any(is.na(x))})
sum(row.has.na)#removed 1 individual here
BARDcasefam<-BARDcasefam[!row.has.na,]#This Fam file is good to go for consent group BARD
BARDcasefam<-mutate(BARDcasefam,V1,GAIN_ID,V3,V4,V5,V6=2)

casestotalfam<-bind_rows(BARDcasefam,BDOcasefam)#Prepared .fam with all of the AA cases
fwrite(casestotalfam, "/home/peter/bipolar_disorder/preQCSteps/caseswithpheno.fam", col.names = F, row.names = F, quote = F, sep ="\t")
caselist<-dplyr::select(casestotalfam,V1,GAIN_ID)
fwrite(caselist,"/home/peter/bipolar_disorder/caselist.txt", col.names = F, row.names = F , quote = F, sep = "\t")
```

6)Prepping dbGaP genotype files so non AA individuals can be excluded:
```{bash}
plink --bfile /home/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype/phg000015.filtered.BARD.plink --bmerge /home/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype/phg000016.filtered.BARD.plink --make-bed --out BARDcases
#BARD Case_genotypes

plink --bfile /home/peter/bipolar_disorder/62470_GAIN_DS-BD_n653/phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype/phg000015.filtered.BDO.plink --bmerge /home/peter/bipolar_disorder/62470_GAIN_DS-BD_n653/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c3.BDO.filtered.matrixfmt.genotype/phg000016.filtered.BDO.plink --make-bed --out BDOcases
#BDO cases genotypes

plink --bfile BDOcases --bmerge /home/peter/bipolar_disorder/62469_GAIN_REL-BD_n841/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c2.BARD.filtered.matrixfmt.genotype/BARDcases --make-bed --out casesbadpheno
#total cases without a good phenotype

plink --bfile casesbadpheno  --keep  /home/peter/bipolar_disorder/preQCSteps/caselist.txt --make-bed --out /home/peter/bipolar_disorder/preQCSteps/AAcases

plink --bed /home/peter/bipolar_disorder/preQCSteps/AAcases.bed --bim /home/peter/bipolar_disorder/preQCSteps/AAcases.bim --fam /home/peter/bipolar_disorder/preQCSteps/caseswithpheno.fam --make-bed --out AAcases4merge #The complete genotype fileset for the cases.  This now just needs to be completed with the cases and merged before QC.

plink --bfile /home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/phg000015.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype/phg000015.filtered.GRU.plink --bmerge /home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype/phg000016.filtered.GRU.plink --make-bed --out GRUcontrols
#Control genotype file

plink --bfile /home/peter/bipolar_disorder/57575_GAIN_GRU-BD_n1767/phg000016.Bipolar.genotype-calls.Affymetrix6.v1.p1.c1.GRU.filtered.matrixfmt.genotype/GRUcontrols --keep /home/peter/bipolar_disorder/preQCSteps/AAcontrolslist.txt  -make-bed --out /home/peter/bipolar_disorder/preQCSteps/AAcontrolsbadpheno

plink --bed /home/peter/bipolar_disorder/preQCSteps/AAcontrolsbadpheno.bed --bim /home/peter/bipolar_disorder/preQCSteps/AAcontrolsbadpheno.bim --fam /home/peter/bipolar_disorder/preQCSteps/AAcontrols.fam --make-bed --out AAcontrols4merge

plink --bfile AAcontrols4merge --bmerge AAcases4merge --make-bed --out aa4QCstart
```